### 2014-03-27 ms_probaweb.tex
paragraph on sampling

Finally, the co-occurrence probability itself could be dependent on ecological
interactions. Direct pairwise interactions such as competition, facilitation and
predation have long been studied for their impact on co-distribution. Second and
higher order interactions (e.g. trophic cascade) could also impact
co-occurrence. There is however currently no general theory on the expected
co-occurrence in complex ecological networks. For instance, we do not know how
far co-occurrence is not-random when going along the chain of indirect
interactions. Berlow(2009) shown previously that almost only first and second
order interactions do matter in food webs, but we don't know for
co-distribution. We neither know what is the sensitiviy to species richness: do
interactions tend to buffer each other? Generalizing knowledge aquired by the
study of small community modiles will require future research.

@@

Part of the problem surrounding the definition of the niche has been
clarified with the distinction between Eltonian and Grinnellian definitions
(ChaseLeibold 2003). The Grinnellian dimension of the niche is the effect of
the environment on the demography of a species, while the Eltonian dimension
is the effect of a species on its environment \emph{sensu lato}.  % SNOB!
The Grinnellian niche is the most intuitive one to apply and is the conceptual
backbone of species distribution models. The Eltonian niche is well known
by community ecologists, but is trickier to turn into predictive models
\parencite{Devictor2010a}. Nonetheless, the development of the niche model
of food web structure (Williams2000) and its parameterization (Williams2010;
Gravel2013) made it more operational, although it has yet to be applied to
more than trophic interactions. %Y'a pas que les food webs dans la vie!

@@

@@

\newpage
%========================================================%
\section*{Acknowledgements}
This is tis a contribution to the Santa Fe Institute working group \emph{Networks over
ecological gradients}. DG received financial support from NSERC and Canada
Research Chair program. TP is supported by a MELS-FQRNT post-doctoral
fellowship.
\newpage

>>>>

Finally, the co-occurrence probability itself could be dependent on ecological
interactions. Direct pairwise interactions such as competition, facilitation and
predation have long been studied for their impact on co-distribution. Second and
higher order interactions (e.g. trophic cascade) could also impact
co-occurrence. There is however currently no general theory on the expected
co-occurrence in complex ecological networks. For instance, we do not know how
far co-occurrence is not-random when going along the chain of indirect
interactions. Berlow(2009) shown previously that almost only first and second
order interactions do matter in ecological networks, but we don't know for
co-distribution. We neither know what is the sensitiviy to species richness: do
interactions tend to buffer each other? Generalizing knowledge aquired by the
study of small community modiles will require future research.

@@

Part of the problem surrounding the definition of the niche has been
clarified with the distinction between Eltonian and Grinnellian definitions
(ChaseLeibold 2003). The Grinnellian dimension of the niche is the effect of
the environment on the demography of a species, while the Eltonian dimension
is the effect of a species on its environment \emph{sensu lato}.  % SNOB!
The Grinnellian niche is the most intuitive one to apply and is the conceptual
backbone of species distribution models. The Eltonian niche is well known
by community ecologists, but is trickier to turn into predictive models
\parencite{Devictor2010a}. Nonetheless, the development of the niche model
of food web structure (Williams2000) and its parameterization (Williams2010;
Gravel2013) made it more operational, although it has yet to be applied to
more than trophic interactions.

@@

\subsection*{Guidance for empirical studies}

Ecological networks are known to be extremely sparse, \emph{i.e.} having far
more absences of interactions that they have interactions. These absences of
interactions, however, can come from different sources. The fact that unequal
sampling at the local scale can affect our understanding of network structure
is well documented \parencite{Martinez1999}. However, in a spatial context,
some interactions may be undocumented because the species involved have
never been observed in co-occurence. Although these are reported as a lack
of interactions, in actuality we cannot make inference about them seeing
that they have never been observed: it is possible that this interaction
may happen should the two species co-occur. A second category of absences
of interactions are those that are reported after multiple observations of
species co-occurence. However, so as to have a confidence in the fact that the
probability of an interaction is low, extensive sampling (that is, several
co-occurences) is needed. Generally, our confidence that the interaction
is indeed impossible will increase when the number of observations of the
species pair. Seeing that this is essentialy a Bernoulli process (what is
the probability that the species will interact given their presence), the
breadth of the confidence interval is expected to saturate after a fixed
number of observations, which can be set as a treshold above which a species
paier has been observed "often enough".

@@

\newpage
%========================================================%
\section*{Acknowledgements}
This is tis a contribution to the Santa Fe Institute working group \emph{Networks over
ecological gradients}. DG received financial support from NSERC and Canada
Research Chair program. TP is supported by a MELS-FQRNT post-doctoral
fellowship. DG, TP, and DBS received financial support from the CIEE.
\newpage
